Capstone Movielens Report

### 2025-12-06 reports/capstone-movielens.site/02.5.UME-model.Rmd
Function described: `train_UMGY_effect`.

[Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_ explains why and how we should use _Penalized least squares_ to improve our predictions. The author also explains that the general idea of penalized regression is to control the total variability of the movie effects[@IDS2_24-1]: 
$$
\sum_{j=1}^{n} \beta_j^2
$$ 
Specifically, instead of minimizing the least squares equation, we minimize an equation that adds a penalty:

>>>>

[Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_ explains why and how we should use _Penalized least squares_ to improve our predictions. The author also explains that the general idea of penalized regression is to control the total variability of the movie effects[@IDS2_24-1]: 
$$
\sum_{j=1}^{n} \beta_j^2
$$ 
Specifically, instead of minimizing the least squares equation, we minimize an equation that adds a penalty:

### 2025-12-06 reports/capstone-movielens.site/05.appendix.Rmd
Function described: `train_UMGY_effect`.

We call this function internally from the functions that compute the _penalized estimates_  as described in [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the *Course Textbook (New Edition)*[@IDS2_24-1].

@@

The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_effect.cv.args.lambda) argument, is other than default (to be more exact, greater than zero).
:::
:::

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_genre_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_genre_effect.cv.args.lambda) argument, is other than default (to be more exact, greater than zero).
:::
:::

@@

The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the *Date General Effect* calculated by this function only when the *Regularization parameter* value, passed in the [lambda](#func.calc_date_general_effect.args.lambda) argument, is non-default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the internally called [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the *Date General Effect* calculated by this function only when the *Regularization parameter* value, passed in the [lambda](#func.calc_date_general_effect.cv.args.lambda) argument, is non-default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the internally called [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
The source code of the [calc_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L71) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L90) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L71) script on _GitHub_.
:::
:::

#### [train_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L42) Function {#func.train_UMGY_effect}
\

Trains (and optionally *regularizes* if the [lambda](#func.train_UMGY_effect.args.lambda) argument value is non-default) the *User+Movie+Genre Effect (UMGYE) Model* using the training sample of the `Movielens` dataset (passed in the [train.sgr](#func.train_UMGY_effect.args.train.sgr) argument), specially prepared to train the *UMGYE Model*.

@@

::: { #func.train_UMGY_effect.args.train.sgr .sidebar }
  - **train.sgr:** Training sample of the `Movielens` dataset with split rows belonging to multiple genres;

@@

::: { #func.train_UMGY_effect.details .sidebar }
##### Details
\
The dataset passed in the [train.sgr](#func.train_UMGY_effect.args.train.sgr) argument is assumed to be the result returned by the [splitGenreRows](#func.splitGenreRows) helper function described above in the *Data Helper Functions /* [Data Processing Functions] section of *This Appendix*.

For example, the [`edx.sgr` Object] is an appropriate dataset as the value of the [train.sgr](#func.train_UMGY_effect.args.train.sgr).

::: {.noteblock data-latex=""}
The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_UMGY_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

Data frame representing the trained (and also *regularized* if [lambda](#func.train_UMGY_effect.args.lambda) arg value other than default) *UMGY Effect Model*.
:::

@@

The source code of the [train_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L42) function is shown below:
```{r eval=FALSE}
```

::: {.noteblock data-latex=""}
The source code of the [train_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L42) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L3) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L42) script on _GitHub_.
:::
:::

@@

Trains (and optionally *regularizes* if the [lambda](#func.train_UMGY_effect.cv.args.lambda) argument value is non-default) the *User+Movie+Genre Effect (UMGYE) Model* with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in section [Section 24.1 *Case study: recommendation systems / Penalized Least Squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_UMGY_effect.cv.args.lambda) argument, is other than default (to be more exact, greater than zero).
:::
:::

@@

Validates the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMGY_effect.cv.args.lambda) argument for the *User+Movie+Genre Effect (UMGYE) Model* in the *regularization process* using *penalized approach* with use of *K-Fold Cross-Validation*.

>>>>

We call this function internally from the functions that compute the _penalized estimates_  as described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the *Course Textbook (New Edition)*[@IDS2_24-1].

@@

The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_effect.cv.args.lambda) argument, is other than default (to be more exact, greater than zero).
:::
:::

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_genre_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_user_movie_genre_effect.cv.args.lambda) argument, is other than default (to be more exact, greater than zero).
:::
:::

@@

The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the *Date General Effect* calculated by this function only when the *Regularization parameter* value, passed in the [lambda](#func.calc_date_general_effect.args.lambda) argument, is non-default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the internally called [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the *Date General Effect* calculated by this function only when the *Regularization parameter* value, passed in the [lambda](#func.calc_date_general_effect.cv.args.lambda) argument, is non-default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the internally called [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.
:::
:::

@@

::: {.noteblock data-latex=""}
The source code of the [calc_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L71) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L3) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L71) script on _GitHub_.
:::
:::

#### [train_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L76) Function {#func.train_UMGY_effect}
\

Trains (and optionally *regularizes* if the [lambda](#func.train_UMGY_effect.args.lambda) argument value is non-default) the *User+Movie+Genre+Year Effect (UMGYE) Model* using the training sample of the `Movielens` dataset passed in the [train_set](#func.train_UMGY_effect.args.train_set) argument.

@@

::: { #func.train_UMGY_effect.args.train_set .sidebar }
  - **train_set:** Training sample of the `Movielens` dataset;

@@

::: { #func.train_UMGY_effect.details .sidebar }
##### Details
\

The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_UMGY_effect.args.lambda) argument, is other than default (to be more exact, greater than zero). The regularization is committed through the *penalized regression* by computing the *penalized mean* performed by the [mean_reg](#func.mean_reg) function described above in the *Common Helper Functions /* [Regularization Functions] section of *this Appendix*.

@@

Data frame representing the trained (and also *regularized* if [lambda](#func.train_UMGY_effect.args.lambda) argument value is non-default) *UMGY Effect Model*.
:::

@@

The source code of the [train_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L76) function is shown below:
```{r eval=FALSE}
train_UMGY_effect <- function(train_set, lambda = 0){
  if (is.na(lambda)) {
    stop("Function: train_UMGY_effect
`lambda` is `NA`")
  }

  train_set |>
    calc_date_general_effect(lambda) |>
    calc_UMGY_effect()
}
```

::: {.noteblock data-latex=""}
The source code of the [train_UMGY_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L76) function is also available in the [Utility Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L3) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L76) script on _GitHub_.
:::
:::

@@

Trains (and optionally *regularizes* if the [lambda](#func.train_UMGY_effect.cv.args.lambda) argument value is non-default) the *User+Movie+Genre+Year Effect (UMGYE) Model* with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

@@

::: {.noteblock data-latex=""}
The *Regularization technique* (described in [Section 24.2 *Penalized least squares*](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#penalized-least-squares) of the _Course Textbook (New Edition)_[@IDS2_24-1]) is applied to the model being trained only when the *Regularization parameter* value, passed in the [lambda](#func.train_UMGY_effect.cv.args.lambda) argument, is other than default (to be more exact, greater than zero).
:::
:::

@@

Validates the *regularization parameter* $\lambda$ passed in the [lambda](#func.regularize.test_lambda.UMGY_effect.cv.args.lambda) argument for the *User+Movie+Genre+Year Effect (UMGYE) Model* in the *regularization process* using *penalized approach* with use of *K-Fold Cross-Validation*.
